---
title: "Project Management"
author: "Ryan Miller"
output: 
  beamer_presentation:
    colortheme: "beaver"
    includes:
      in_header: pres_header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Organizational Structure

\includegraphics[width=1\textwidth]{fig/org.png}

## Data Science "pipelines" and "life-cycles"

- The day-to-day data science tasks take place within the organizational framework
- While "the data science process" is often discussed, its definitions vary
    - [You've seen some examples at this link](https://remiller1450.github.io/s395s20/ds_proj_pipe.pdf)
  
## CRISP-DM

We'll adopt the Cross-Industry Standard Process for Data Mining (CRISP-DM) as our model:

\begin{center}
\includegraphics[width=0.67\textwidth]{fig/crisp_process.png}
\end{center}

## Phase 1 - Business Understanding

Tasks:

1) Gather necessary [background information](https://www.ibm.com/support/knowledgecenter/SS3RA7_15.0.0/com.ibm.spss.crispdm.help/crisp_business_criteria.htm)
2) Document specific [specific objectives](https://www.ibm.com/support/knowledgecenter/SS3RA7_15.0.0/com.ibm.spss.crispdm.help/crisp_business_objectives.htm)
3) Determine [success criteria](https://www.ibm.com/support/knowledgecenter/SS3RA7_15.0.0/com.ibm.spss.crispdm.help/crisp_success_criteria_business.htm) for the project

Each of these tasks should be undertaken in coordination with the client 

## Phase 1 - Objective vs. Subjective Success Criteria

- You may choose to have a mixture of objective and subjective success criteria
    - Objective = "Increase the time visitors spend on the landing page by 10%"
    - Subjective = "Identify customer clusters for targeted marketing"
    
## Phase 2 - Data Understanding

Tasks:

1) [Describing](https://www.ibm.com/support/knowledgecenter/SS3RA7_15.0.0/com.ibm.spss.crispdm.help/crisp_describe_data.htm) the data
2) [Exploring](https://www.ibm.com/support/knowledgecenter/SS3RA7_15.0.0/com.ibm.spss.crispdm.help/crisp_explore_data.htm) the data
3) [Verifying data quality](https://www.ibm.com/support/knowledgecenter/SS3RA7_15.0.0/com.ibm.spss.crispdm.help/crisp_verify_data_quality.htm)

These tasks should be carried out at the team level (and cross-referenced with the principal and client if necessary)

## Phase 3 - Data Preparation

Tasks:

- Merging/joining (ie: `left_join`)
- Selecting relevant subsets (ie: `filter`)
- Aggregating records (ie: `group_by` and `summarize`)
- Deriving new attributes (ie: `mutate`)
- Handling missing data (ie: `complete.cases` or `knnImput`/`rfImpute`)

These tasks should be carried out at the team level and cross-referenced by the principal (they are seldom relevant to the client at this point)

## Phase 4 - Modeling 

Tasks:

- [Selecting](https://www.ibm.com/support/knowledgecenter/SS3RA7_15.0.0/com.ibm.spss.crispdm.help/crisp_select_model_technique.htm) a model
- [Evaluating the "goodness"](https://www.ibm.com/support/knowledgecenter/SS3RA7_15.0.0/com.ibm.spss.crispdm.help/crisp_generate_test_design.htm) of a model
- [Building](https://www.ibm.com/support/knowledgecenter/SS3RA7_15.0.0/com.ibm.spss.crispdm.help/crisp_build_model.htm) the model
- Note: you may replace "model" with "product" in some applications

This phase is highly non-linear, it should be carried out at the team level and cross-referenced by the principal (and possibly with the client depending on their level of technical proficiency)

## Phase 5 - Evaluation 

Tasks:

1) Consider your model/product in regards to the business success criteria you came up with in Phase 1
2) Formalize your findings

These tasks should be undertaken in coordination with the client and principal

## Phase 6 - Deployment

Tasks:

1) Deliver your model/product to the client
2) Complete wrap-up tasks (ie: technical report, etc.)

## Review #1

For the following scenario determine:

A) Which phase the described actions fall under
B) Where you'd go next (and why)

> A project is using medical records to build a model to predict A1c levels using more readily available measures such as blood pressure, age, weight, and waist circumference.  Using the `is.na` function in `R`, it is discovered that 88% of the available records do not have an A1c measurement.


## Review #1 - Possible Answers

- This scenario occurred during the **Data Understanding** phase
- It is possible that the 12% of medical records with A1c provides a reasonable dataset, in this case the next step would be Data Preparation (filtering out the missing data and preparing the other variables)
- It is also possible that relying on only 12% of the available records is infeasible or will induce bias into the analysis, in this case the next step would be Business Understanding (reviewing the original goals and making adjustments)

## Review #2

For the following scenario determine:

A) Which phase the described actions fall under
B) Where you'd go next (and why)

> In the aforementioned project you fit a linear regression model containing several variables, you receive the following model diagnostics from your software

```{r, fig.height = 4}
set.seed(1234)

x <- rnorm(500)
y4 <- (.1 * x) + rnorm(500)
y1 <- 2 * x + rnorm(500)
y2 <- 2 * x + - (.5 * x^2) + rnorm(500)
y3 <- exp(.5 * x + rnorm(500))

# put data into dataframe to organize results
df <- data.frame(x, y1, y2, y3, y4)

# run regressions
regs <- lapply(df[-1], function(y) lm(y ~ x, data=df))

par(mfrow = c(1,2))
plot(regs$y2, which = 1:2)
```


## Review #2 - Possible Answers

- This scenario occurred during the **Modeling phase**
- Linear regression doesn't appear to be an appropriate model based upon these diagnostics.  It seems that a quadratic effect is being missed (hence the large negative residuals for high/low fitted values). The QQ-plot also calls into question whether the residuals are normally distributed (not a disaster for model fitting, but a problem for statistical inference)
    - For these reasons the logical next step is to return to the Data Preparation phase and explore variable transformations that might address these issues

## Review #3 

For the following scenario determine:

A) Which phase the described actions fall under
B) Where you'd go next (and why)

> After revising the linear regression model in the previous example, a final model is chosen and is applied a "test set" of 100 new records that occurred after the original dataset was finalized. The model predicts A1c within 10% of the actual value for 86% of these new records.

## Review #3 - Possible Answers

- This scenario occurred in the **Evaluation phase**
- Where to go next *depends upon the project's business goals*.  If predicting A1c within 10% of the actual value for 86% of cases satisfies the previously established goals, the model is suitable for the Deployment phase.  Otherwise the project might need to return to square one.

